An Algorithm For Spatial Mapping Using a Hexagon Tilegram, With Application to Australian Maps

### 2019-07-29 algorithmRjournal/algorithmRjournal.tex
Tasmania

Spatial distributions have utilised alternative representations of
geography for many years. In modern times interactivity and animation
have begun to play a larger role, as alternative representations have
been popularised by online news sites, with a focus on public
consumption. Applications are increasingly widespread, especially in the
areas of disease mapping, and election results.

\hypertarget{motivation}{%
\subsection{Motivation}\label{motivation}}

Spatial distributions of people and human related statistics can be
misrepresented in geographic maps. Populations congregate around major
cities and vertical living is increasingly common. Population statistics
often requires dividing people into smaller, measurable areas.
Government bodies such as the Australian Bureau of Statistics, or
intentionally segregated organisations, like the Australian Electoral
Commission hold the responsibility of the division. Australia is an
example of strong urbanisation, the rural areas are often sparsely
populated in comparison to the urban centres. To divide the population
equally, the square meters of the geographic areas differ dramatically
in size.

Alternative mapping methods allow increased understanding of the spatial
distribution of a variable across the population. Alternative maps allow
the focus to be placed on the distribution of the statistics between the
groups of areas.

Extremely small geographic areas are lost on large geographic maps,
small inner areas of Sydney or Melbourne are not easily compared at a
high level. Cartograms place importance on the statistic of interest,
allowing distorted map space on the display to represent differences in
the statistic. These maps rely on the statistic of interest to determine
the layout, and for Australia, often fail to preserve a recognisable
view. Tilegrams allow preservation of spatial relationships as they
decrease the emphasis on the amount of geographic area considered to be
interesting. These maps focus on the relationship between neighbours, in
considering aggregated statistics of heterogeneous regions.

Our case is an extension of these concepts. Extending the tilegram to
Australian applications required preserving the spatial relationships.
It emphasises the capital cities as population hubs, and recognises the
populations distributed across the large rural areas.

@@

Our solution operates on a set of polygons. There are parameters used in
the process that may be provided, or will be automatically derived. All
necessary functions are exported, with a main function
\texttt{create\_hexmap} used to step through these automatically.

\hypertarget{parameters}{%
\subsection{Parameters}\label{parameters}}

The \texttt{create\_hexmap} function requires several parameters, if
they are not provided, the information will be derived from the polygon
set used.

@@

\begin{tabular}{l|l}
\hline
parameter & description\\
\hline
shp & an Rdata object containing the polygon information\\
\hline
shp\_path & character string location of the shape file that contains the polygon information\\
\hline
sf\_id & name of a unique column that distinguishes areas\\
\hline
capital\_cities & a data frame of reference locations used to allocate hexagons\\
\hline
\end{tabular}

@@

The polygon set of Local Government Areas in Tasmania in 2016 is stored
in the sugarbag package data as \texttt{tas\_lga}. A single column of
the data set must be used as a unique identifier of areas. In this case,
the unique LGA codes associated with each LGA have been used, their
names could have also been used.

The centre of the population hubs are used to distribute areas around
the closest capital city.

@@

\begin{tabular}{l|l}
\hline
parameter & description\\
\hline
buffer\_dist & a float value for distance in degrees to extend beyond the - geometry provided\\
\hline
hex\_size & a float value in degrees for the diameter of the hexagons\\
\hline
hex\_filter & amount of hexagons around centroid to consider for allocation\\
\hline
width & the angle used to filter the grid points around a centroid\\
\hline
\end{tabular}

@@

When utilising individual \texttt{sugarbag} functions, we recommend the
following approach:

\hypertarget{polygon-information}{%
\subsection{Polygon Information}\label{polygon-information}}

Begin with a set of polygons. The set should be provided as an
\texttt{sf} object, this is a data frame containing a \texttt{geometry}
column. The \texttt{read\_shape} function can assist in creating this
object. We will use the Local Government Areas of Tasmania, sourced from
the Australian Bureau of Statistics.

\begin{Schunk}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-3-1} \caption[lga Local Government Areas and derived centroids]{lga Local Government Areas and derived centroids.}\label{fig:unnamed-chunk-3}
\end{figure}
\end{Schunk}

@@

Each new hexagon will need to tessellate, a grid is created to ensure a
close point is found, that will allow all areas to tessellate, this is
inspired by tilegrams. The grid of possible hexagon centroids can be
made using the \texttt{create\_grid} function. The grid creation takes
several steps. It requires the centroids, the hexagon size and the
buffer distance.

@@

\includegraphics{algorithmRjournal_files/figure-latex/plot_grid-1} \end{Schunk}

@@

\includegraphics{algorithmRjournal_files/figure-latex/plot_shift-1} \end{Schunk}

@@

@@

Not all of the grid points will be used, especially if islands impact
the overall space covered. To filter the grid for appropriate points for
allocation, the \texttt{create\_buffer} function is called from
\texttt{create\_grid}. It finds the amount of columns and rows needed to
best capture the set of centroids.

The rows and columns are divided into 20 groups. The amount of rows in
each latitude group and the amount of columns in each longitude group
are then used as the width of manual rolling windows. This will tailor
the available points to the areas most likely to be used. It helps
reduce amount of time taken.

The specific rows and columns in the rolling windows are defined.The
first rolling window function finds the minimum and maximum centroid
values for the sliding groups of longitude columns and the groups of
latitude rows.

The second rolling window function finds the average of the rolling
minimum and rolling maximum centroid values, for the longitude columns
and latitude rows.

@@

@@

\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-16-1} \end{Schunk}

@@

\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-19-1} \end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
# Create hexagon location grid
grid <- create_grid(centroids = centroids,
    hex_size = 0.25,
    buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "LGA_CODE16",
  # same column used in create_centroids
  hex_size = 0.25,
  # same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities %>% filter(points == "Hobart"),
  width = 30,
  verbose = FALSE
)
\end{Sinput}
\end{Schunk}

@@

\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-20-1} \end{Schunk}

@@

\address{%
Stephanie Kobakian\\
Queensland Univeristy of Technology\\
\\
}
\href{mailto:stephanie.kobakian@monash.edu}{\nolinkurl{stephanie.kobakian@monash.edu}}

>>>>

Spatial distributions have utilised alternative representations of
geography for many years. In modern times, interactivity and animation
have begun to play a larger role, as alternative representations have
been popularised by online news sites, and atlas websites with a focus
on public consumption. Applications are increasingly widespread,
especially in the areas of disease mapping, and election results.

Choropleths are the current best practice approach for presenting
geographical data as maps highlight the geographic patterns in
geospatially related statistics {[}@SAMGIS{]}. The land on the map space
is divided to separate each area, these boundaries are usually
administrative. The areas are filled with colour to represent the value
of the statistic for the area {[}@EI{]}.

In Australia, government bodies such as the Australian Bureau of
Statistics, or the Australian Electoral Commission hold the
responsibility for the division of areas. These boundaries are often
adjusted as the population in areas increase. Australian people are
increasingly congregating around major cities, the rural areas are often
sparsely populated in comparison to the urban centres. The division of
the population into equal population areas results in dramatically
different square meterage of the geographic areas. This results in an
unequal presentation of the statistic in each area, this can
misrepresent the spatial distributions of the human related statistics
in geographic maps.

Beginning with the geography, cartograms apply a transformation to place
importance on the statistic of interest. They result in a distortion of
the map space to represent differences in the statistic. These displays
rely on the statistic of interest to determine the layout, however for
Australia, many cartograms fail to preserve a recognisable display.
Contiguous cartograms change the shape of an area, preserving
neighbouring relationships. Non-contiguous cartograms maintain the shape
of each area, but lose the connection to neighbours.

Alternative maps shift the focus from land area, to the value of the
statistics in a group of areas. Alternative mapping methods allow
increased understanding of the spatial distribution of a variable across
the population, by fairly representing each administrative area.

Tilegrams are an extension of Dorling cartograms, using one simple shape
to represnt each area. This allows preservation of spatial relationships
and decreases the emphasis from the amount of geographic area. These
maps focus on the relationship between neighbours attempting to preserve
connections, and disregard the unique shapes of the administrative
boundaries.

The sugarbag package provides a new method to implement hexagon
tilegrams. Extending the tilegram to Australian applications required
preserving the spatial relationships. It emphasises the capital cities
as population hubs, and recognises the distances between locations of
populations distributed across the large rural areas.

@@

This solution operates on a set of \texttt{sf} {[}@sf{]} polygons.

There are four necessary tasks to create a hexagon tilegram. These steps
can be executed by the main function, \texttt{create\_hexmap}, or can be
applied separately for more flexibility.

They perform the following tasks: - Create the set of centroids to
allocate - Create the set of hexagons to use - Allocate each centroid to
an available hexagon - Transform the data for plotting

There are parameters used in the process that may be provided, or will
be automatically derived.

\hypertarget{use-the-sugarbag-package}{%
\subsubsection{Use the sugarbag
package}\label{use-the-sugarbag-package}}

\hypertarget{parameters}{%
\subsubsection{Parameters}\label{parameters}}

The \texttt{create\_hexmap} function requires several parameters, if
they are not provided, the information will be derived from the polygon
set used. If users already have a grid and set of centroids they may
choose to only use the \texttt{allocate} function.

@@

\begin{tabular}{l|l}
\hline
parameter & description\\
\hline
shp & an sf object containing the polygon information\\
\hline
sf\_id & name of a column that distinguishes unique areas\\
\hline
focal\_points & a data frame of reference locations used to allocate hexagons\\
\hline
\end{tabular}

@@

\hypertarget{example-data}{%
\subsubsection{Example data}\label{example-data}}

The polygon set of Local Government Areas of Tasmania in 2016 is
provided with the sugarbag package as \texttt{tas\_lga}. A single column
of the data set must be used as a unique identifier of areas. In this
case, the unique LGA codes associated with each LGA have been used.

The longitude and latitude centre of the capital cities of Australia are
used to allocate areas around the closest capital city.

@@

\begin{tabular}{l|l}
\hline
parameter & description\\
\hline
buffer\_dist & a float value for distance in degrees to extend beyond the geometry provided\\
\hline
hex\_size & a float value in degrees for the diameter of the hexagons\\
\hline
hex\_filter & amount of hexagons around centroid to consider for allocation\\
\hline
width & the angle used to filter the grid points around a centroid\\
\hline
\end{tabular}

@@

\hypertarget{create-the-set-of-centroid-points}{%
\subsection{Create the set of centroid
points}\label{create-the-set-of-centroid-points}}

When utilising individual \texttt{sugarbag} functions, we recommend the
following approach: Begin with a set of polygons. The set should be
provided as an \texttt{sf} object, this is a data frame containing a
\texttt{geometry} column. \footnote{The \texttt{read\_shape} function
  can assist in creating this object.} We will use the Local Government
Areas of Tasmania, downloaded from the Australian Bureau of Statistics.

The centroids can be derived from the set of polygons:

\begin{Schunk}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/end_cents-1} \caption[Local Government Areas and derived centroids]{Local Government Areas and derived centroids.}\label{fig:end_cents}
\end{figure}
\end{Schunk}

@@

A tilegram allows areas to be allocated to a tellesated set of tiles.
The grid is created to ensure a close point is found, that ensures
tessellation with other previously allocated points.

The grid of possible hexagon centroids can be made using the
\texttt{create\_grid} function. The grid creation requires several
steps. It uses the centroids, the hexagon size and the buffer distance.

@@

@@

@@

\begin{Schunk}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-3-1} \caption[Grid points to create a tilegram]{Grid points to create a tilegram.}\label{fig:unnamed-chunk-3}
\end{figure}
\end{Schunk}

@@

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate points for
allocation, the \texttt{create\_buffer} function is called from
\texttt{create\_grid}. It finds the grid points needed to best capture
the set of centroids.

The rows and columns are divided into 20 groups. The amount of rows in
each latitude group and the amount of columns in each longitude group
are then used as the width of rolling windows. This will tailor the
available grid points to the most likely to be used point. It also helps
reduce amount of time taken, as it decreases the amount of points
considered.

The first rolling window function finds the minimum and maximum centroid
values for the sliding groups of longitude columns and the groups of
latitude rows.

The second rolling window function finds the average of the rolling
centroid values minimum and maximum, for the longitude columns and
latitude rows.

@@

\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-10-1} \end{Schunk}

@@

@@

\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-17-1} \end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
# Create hexagon location grid
grid <- create_grid(centroids = centroids,
    hex_size = 0.3,
    buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "LGA_CODE16",
  # same column used in create_centroids
  hex_size = 0.3,
  # same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities %>% filter(points == "Hobart"),
  width = 35,
  verbose = FALSE
)
\end{Sinput}
\end{Schunk}

@@

\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-18-1} \end{Schunk}

@@

\address{%
Stephanie Kobakian\\
Queensland University of Technology\\
\\
}
\href{mailto:stephanie.kobakian@monash.edu}{\nolinkurl{stephanie.kobakian@monash.edu}}
